---
title: "R Notebook"
output: html_notebook
---

test test text text

I have downloaded csv files of closing prices of the following assets:

Crypto Assets:
1. Bitcoin (BTC)
2. Ethereum (ETH)
3. Tether (USDT)
4. USD Coin (USDC)
5. Binance Coin (BNB)

Traditional Assets
1. Vanguard 500 Index Fund (VOO) (Equities index)
2. Invesco QQQ Trust (QQQ) (Equities index)
3. Vanguard Total Bond Market Index Fund (VBTLX) (Bonds index)
4. abrdn Physical Gold Shares ETF (SGOL) (Metals)
5. Fidelity ZERO Large Cap Index Fund (FNILX) (Equity index large cap)

I'll load these CSV files into a combined dataframe
```{r}
library(tidyverse)

BTC <- read.csv("BTC-USD.csv")
BTC <- tail(BTC[,5],250)

ETH <- read.csv("ETH-USD.csv")
ETH <- tail(ETH[,5],250)

USDT <- read.csv("USDT-USD.csv")
USDT <- tail(USDT[,5],250)

USDC <- read.csv("USDC-USD.csv")
USDC <- tail(USDC[,5],250)

BNB <- read.csv("BNB-USD.csv")
BNB <- tail(BNB[,5],250)

VOO <- read.csv("VOO.csv")
VOO <- tail(VOO[,5],250)

QQQ <- read.csv("QQQ.csv")
QQQ <- tail(QQQ[,5],250)

VBTLX <- read.csv("VBTLX.csv")
VBTLX <- tail(VBTLX[,5],250)

SGOL <- read.csv("SGOL.csv")
SGOL <- tail(SGOL[,5],250)

FNILX <- read.csv("FNILX.csv")
Date<- tail(FNILX[,1], 250)
FNILX <- tail(FNILX[,5],250)


data_combine <- data.frame(Date,BTC,ETH,USDT,USDC,BNB,VOO,QQQ,VBTLX,SGOL,FNILX)

```

I'm  converting my DF into a time series object, and plotting the series

```{r}
require(xts)
library(xts)

data_combine[data_combine == "null"] <- ""
#View(data_combine)

df <- sapply(data_combine, as.numeric)
str(df)
#df1 <- null.omit(df)
#df2 <- na.omit(df1)
df3 <- data.frame(date=Date,df)


df3$date = as.POSIXct(df3$date,format="%Y-%m-%d", tz = "")
df3.xts <- xts(df3[,-1], order.by=df3[,1])

plot(df3.xts)
#plot(df3.xts[,1:5])
#plot(df3.xts[,6:10])
```

Plotting returns
```{r}

#log return
return = diff (log(df3.xts)) # Log return calculation
return = return [-1] # removing the first empty observation, received after return calculation
summary (return)
plot(return)
```


```{r}
library(PerformanceAnalytics)
return2 = Return.calculate(df3.xts, method="log")

return3 = diff (df3.xts) # differences
# the sane as 
return4 = Return.calculate(df3.xts, method="difference")
#daily gain/loss (%)
return5 = diff(df3.xts)/df3.xts[-length(df3.xts)]

plot(return5)
```


```{r}
#Correlation #Chapter 2: Session 2
cor(df3.xts)
```
We can notice that Bitcoin and Etherium is highly correlated to each other in a positive fasion.
There is some positive correlation between BTC and Tether coin.

Among traditional assets: VOO is positively correlated to QQQ and FNILX, as they are both equity index funds.
Correlation amon Crypto and traditional assets are very low, and often even negative.

## Part 2

```{r}
##ECONOMIC FACTOR MODEL
Traditional_Assets <- df3.xts[,7:11]
Crypto <- df3.xts[,1:6]
```

#### Factor Analysis

```{r}
## FACTOR Analysis

Traditional_Assets2 = as.matrix(Traditional_Assets)   
Crypto2 = as.matrix(Crypto) 
Indices = as.data.frame(diff(log(Traditional_Assets2)))  # log difference
names(Indices)[1]="Traditional_Assets" 
Cryptocurrency = as.data.frame(diff(log(Crypto2))) #calculte log difference 
names(Cryptocurrency)[1] = "Cryptocurrencies"
Indices_Cryptocurrency = cbind(Indices,Cryptocurrency) #View(Indices_Cryptocurrency)
berndt = as.matrix(df3.xts[,-1])
arFit = ar(cbind(Indices,Cryptocurrency)) # autoregressive model - by default selecting the complexity by AIC. Here it is AR(5)



```

```{r}

res = arFit$resid[4:1446,] # residuals of the AR (5) modle (the interpretations of the residuals is that they are unexpected shocks)
lmfit = lm(berndt[4:1446,2:9]~res[,1]+res[,2]) #fit a regression Y is a set of teh stock returns and X1 and X2 - residuals from the AR modles
slmfit = summary(lmfit) #summary of the regression
length(berndt[,1]) #dim(berndt[,1])
```

```{r}

beta_CPI = lmfit$coef[2,] # extract all world indices betas (b1)
beta_IP = lmfit$coef[3,] # extract all Cruptocurrences betas (b2)
```

####Determining the Number of Factors to Extract
### Scree plot

```{r}
library(Momocs)
library(psych)
scree(df3.xts)

#Parallel Analysis
fa.parallel (df3.xts) #

```

Based on the scree plot of Principal component and factor analysis we can see that 3 factors will be adequate to explain the maximum variation in the data.

```{r}
par(mfrow=c(1,3)) # building three graphs in a row
#barplot(rsq,horiz=T,names=names(beta_CPI),main="R squared") #Creates a bar plot with vertical or horizontal bars.
barplot(beta_CPI,hori=T,main="beta CPI") #Creates a bar plot with vertical or horizontal bars.
barplot(beta_IP,hori=T,main="beta IP") #Creates a bar plot with vertical or horizontal bars.

```


```{r}

fa_none = factanal(df3.xts[,2:10],4,rotation="none") # 4 factor model without rotation
print(fa_none,cutoff=0.1) # By convention, any loading with an absolute value less than the parameter cutoff is not printed, and the default value of cutoff is 0.1
```

Since the p-value is very low, we can go ahead and reject the hypothesis that the model fits the data well.

### VARIMAX rotation

```{r}
fa_vari = factanal(df3.xts[,2:10],4,rotation="varimax") #factor model with rotation

print(fa_vari,cutoff=0.1,sort=T)
print(fa_vari,cutoff=0.1)
sum(fa_vari$loadings[,1]^2)
B=fa_vari$loadings[,] # factor loadings
B
```

We see that the p-value is low here too using varimax rotation
Cumulative variance for 4 factors is just over 70%


### Estimation factor model
```{r}
df3.xts <- subset(df3.xts, select = -c(Date) )
factor.model <- fa(df3.xts, nfactors = 3, fm="ols", max.iter = 100, rotate = "oblimin")
factor.model
```
100 iterations or tries are used to discover the "optimal" answer while estimating a factor model with three elements using OLS. 
Rotate - we employ oblimin rotation in order to allow elements to correlate.


```{r}
fa.diagram(factor.model) 
```

# Communality 
```{r}
factor.model$communality
```
Common variance
#Eeigenvalues
```{r}
factor.model$e.values
```

#Percentage of Variance Accounted For
```{r}
100*factor.model$e.values/length(factor.model$e.values)
```

```{r}
print(factor.model$loadings, cutoff=0, digits=3)
print(factor.model$Structure, cutoff=0, digits=3)
```